A Reduced Order Model for Permeability Fields Using Singular Value Decomposition (SVD)

### 2018-04-12 Papers/AppliedMaths2018/Stochastic_Upscaling_UK_AMM_2018.tex
updates

Upscaling techniques which best preserve statistical properties and flow dynamics behaviour of the high resolution domain give the best representation of heterogeneous properties. There are several upscaling techniques which \red{fall} under different categories \cite{Hasting_2001, Renard_1997, Szymkiewicz_2013}. Traditional upscaling techniques are deterministic in nature, however over the last decade, techniques classed as stochastic have received great attention from academic and industrial porous media communities worldwide \cite{Guilleminot_2012, Ravalec-Dupin_2010, Verwoerd_2009}. One of the reasons for this is because stochastic upscaling techniques are robust enough to handle multiphase flow in porous media and they are able to address scaling and uncertainity problems.

{\it Upscaling} is a term commonly used in a few industries, and it is often referred throughout the applied mathematics and engineering communities as a form of model order reduction (MOR), or the output (\ie Reduced Order Model, ROM). Depending on the context in which MOR is being used, it can simply be referred to as dimensionality reduction. The definition of MOR also depends on the context it is being used \cite{Schilders2008}, but it suffices to state that MOR captures the essential features of a system, structure or parameter. This implies that an original system, structure or parameter exists with more detailed information. Fig.~\ref{fig:IllustrationMOR} presents a graphical intepretation of the MOR concept. It shows that even with the reduced number of facets, the graphic on the right of this illustration is recognized as that of a rabbit.

>>>>

Upscaling techniques which best preserve statistical properties and flow dynamics behaviour of the high resolution domain give the best representation of heterogeneous properties. There are several upscaling techniques which fall under different categories \cite{Hasting_2001, Renard_1997, Szymkiewicz_2013}. Traditional upscaling techniques are deterministic in nature, however over the last decade, techniques classed as stochastic have received great attention from academic and industrial porous media communities worldwide \cite{Guilleminot_2012, Ravalec-Dupin_2010, Verwoerd_2009}. One of the reasons for this is because stochastic upscaling techniques are robust enough to handle multiphase flow in porous media and they are able to address scaling and uncertainity problems.

{\it Upscaling} is a term commonly used in the oil and gas sector and is often referred, throughout the applied mathematics and engineering communities, as a form of model order reduction (or reduced order models, MOR/ROM). Depending on the context in which MOR is used, it can simply be referred as dimensionality reduction. The definition of MOR also depends on the context it is being used, but it suffices to state that MOR captures the essential features of a system, structure or parameter \cite{Schilders2008}. This implies that an original system exists with more detailed information. Figure~\ref{fig:IllustrationMOR} presents a graphical intepretation of the MOR concept in which it shows that even with the reduced number of facets, the graphic on the rhs of this illustration is recognised as that of a rabbit.
